LaMachine: A meta-distribution for NLP software

### 2019-03-21 docs/lamachine_clarin2018.tex
[paper] processed mostly language remarks by reviewer 1

Software is a broad but well-ingrained notion in society nowadays, referring to any form of computer program. This can
manifest in various forms, whether it is an app on a phone, a graphical desktop application, a command line tool, or a
web-based application. These different \emph{interfaces} generally address different \emph{audiences}; the data
scientist will feel at home at the command line and in scripting environments and uses these fairly low-level
interfaces, whereas researchers in the humanities demand higher-level graphical interfaces. There is often a power trade-off
between lower and higher-level interfaces, with the former providing maximum flexibility at the cost of a steeper
learning curve, technical ability, and a do-it-yourself mentality. Higher level interfaces, on the other
hand, expose certain functionality of the software in an easy and accessible way, but in doing so often can not expose
the full power of the software. The cost trade-off is also apparant in the construction of the interfaces, where
high-level interfaces are typically far more costly to build.

@@

LaMachine is an open-source NLP software distribution.  LaMachine facilitates the installation, distribution and configuration of software. It does not fork, modify or appropriate the participating software in any way, nor does it provide a hosting place or repository for software.
We classify LaMachine as a \emph{meta distribution} as it can be installed in various contexts. The heart of LaMachine consists of a
set of machine-parsable instructions on how to obtain, build (e.g. compile from source), install and configure software.
%iris:  eerst algemene uitleg over lamachine- dit is veel te gedetaileerd op dit punt-
%These are implemented using Ansible\footnote{\url{https://www.ansible.com}; a popular open-source software provisioning,
%configuration and deployment tool by Red Hat}.
This is notably different from the more
classical notion of Linux distributions, which generally provide their own repositories with (often binary) software
packages. LaMachine builds on this already established infrastructure by taking these repositories as a foundation where
possible. Similarly, as implied in point five above, there are different programming-language-specific ecosystems
providing their own repositories, such as the Python Package Index\footnote{\url{https://pypi.org}} for Python,
CRAN\footnote{\url{https://cran.r-project.org/}} for R, CPAN\footnote{\url{https://www.cpan.org}} for Perl, Maven
Central\footnote{\url{https://search.maven.org}} for Java.  LaMachine relies on those to obtain and install software. In doing so, we compel participating software projects
to adhere to well-established distribution standards and ensure the software is more sustainable towards the future
\cite{softwarequality}. Moreover, we ensure that LaMachine never becomes a prerequisite for the software but merely a
courtesy or convenience, and attempt to limit any amount of duplication in packaging and distribution efforts.

@@

    \item The software must have a recognised (i.e. OSI-approved) open-source license.
        Proprietary (closed-source) software is explicitly excluded.
    \item The source
        code must be hosted in a public version controlled repository.\footnote{e.g. Github, Gitlab, Bitbucket, provided
        the respository is public}
    \item There must be a build process to compile the source, if applicable, and install the program or library.
    \item There must be some kind of release protocol (adhering to semantic versioning) that publishes software using the proper
technology-specific channels. We will elaborate on this later.
    \item All software that is incorporated in LaMachine must bear at least some relevance to the field of Natural Language Processing.
    \item Participating software must be actively maintained (i.e. not outdated or abandoned) and not place any demands on outdated
        dependencies. %verwerkt <-- maarten help even met formuleren hier, ik bedoel dat zoiets als Pattern dus helaas niet mee kan doen omdat het niet meer onderhouden wordt }only maintained/uptodate date software that can compile using the current version of compilers and dependencies/libraries
\end{enumerate}

@@

     \item \textbf{KafNafParserPy}\footnote{\url{https://github.com/cltl/KafNafParserPy}} - A python module to parse NAF files.

@@

The software metadata discussed here is of a generic and practical nature, it most importantly contains information regarding the version, source,
licensing, and authors of the software. The underlying principle is to obtain sofware metadata from as close to the
source as possible. This can be contrasted with other metadata efforts, also within CLARIN, to establish a more
centralised metadata store\footnote{such as CLARIN's CMDI Component Registry;
\url{https://www.clarin.eu/content/component-metadata}} with more specialised and
research-oriented metadata. There is merit to both approaches as they serve distinct functions. But in order to best
serve everybody's different metadata needs, these bottom-up and top-down processes need to come to some kind of
synthesis, e.g. by having component registries do automatic harvesting of CodeMeta (or other) metadata, similar to what LaMachine
does.

>>>>

Software is a broad but well-ingrained notion in society nowadays, referring to any form of computer program. This can
manifest in various forms, whether it is an app on a phone, a graphical desktop application, a command line tool, or a
web-based application. These different \emph{interfaces} generally address different \emph{audiences}; the data
scientist will feel at home at the command line and in scripting environments and uses these fairly low-level
interfaces, whereas researchers in the humanities demand higher-level graphical interfaces. There is often a power trade-off
between lower and higher-level interfaces, with the former providing maximum flexibility at the cost of a steeper
learning curve, technical ability, and a do-it-yourself mentality. Higher level interfaces, on the other
hand, expose certain functionality of the software in an easy and accessible way, but in doing so often can not expose
the full power of the software. The cost trade-off is also apparent in the construction of the interfaces, where
high-level interfaces are typically far more costly to build.

@@

LaMachine is an open-source NLP software distribution.  LaMachine facilitates the installation, distribution and configuration of software. It does not fork, modify or appropriate the participating software in any way, nor does it provide a hosting place or repository for software.
We classify LaMachine as a \emph{meta distribution} as it can be installed in various contexts. The heart of LaMachine consists of a
set of machine-parsable instructions on how to obtain, build (e.g. compile from source), install and configure software.
%iris:  eerst algemene uitleg over lamachine- dit is veel te gedetaileerd op dit punt-
%These are implemented using Ansible\footnote{\url{https://www.ansible.com}; a popular open-source software provisioning,
%configuration and deployment tool by Red Hat}.
This is notably different from the more
classical notion of Linux distributions, which generally provide their own repositories with (often binary) software
packages. LaMachine builds on this already established infrastructure by taking these repositories as a foundation where
possible. Similarly, as implied in point five above, there are different programming-language-specific ecosystems
providing their own repositories, such as the Python Package Index\footnote{\url{https://pypi.org}} for Python,
CRAN\footnote{\url{https://cran.r-project.org/}} for R, CPAN\footnote{\url{https://www.cpan.org}} for Perl, Maven
Central\footnote{\url{https://search.maven.org}} for Java.  LaMachine relies on those to obtain and install software. In doing so, we compel participating software projects
to adhere to well-established distribution standards and ensure the software is more sustainable
\cite{softwarequality}. Moreover, we ensure that LaMachine never becomes a prerequisite for the software but merely a
courtesy or convenience, and attempt to limit any amount of duplication in packaging and distribution efforts.

@@

    \item The software must have a recognised (i.e. OSI-approved) open-source license.
        Proprietary (closed-source) software is explicitly excluded.
    \item The source
        code must be hosted in a public version controlled repository.\footnote{e.g. Github, Gitlab, Bitbucket, provided
        the repository is public.}
    \item There must be a build process to compile the source, if applicable, and install the program or library.
    \item There must be some kind of release protocol (adhering to semantic versioning) that publishes software using the proper
technology-specific channels. We will elaborate on this later.
    \item All software that is incorporated in LaMachine must bear at least some relevance to the field of Natural Language Processing.
    \item Participating software must be actively maintained (i.e. not outdated or abandoned) and not place any demands on outdated
        dependencies. %verwerkt <-- maarten help even met formuleren hier, ik bedoel dat zoiets als Pattern dus helaas niet mee kan doen omdat het niet meer onderhouden wordt }only maintained/uptodate date software that can compile using the current version of compilers and dependencies/libraries
\end{enumerate}

@@

     \item \textbf{KafNafParserPy}\footnote{\url{https://github.com/cltl/KafNafParserPy}} - A python module to parse NAF
         files, another format for linguistic annotation.
     \item \textbf{NafFoLiApy}\footnote{\url{https://github.com/cltl/NafFoLiApy}} - Converters between FoLiA and NAF.

@@

The software metadata discussed here is of a generic and practical nature, it most importantly contains information regarding the version, source,
licensing, and authors of the software. The underlying principle is to obtain sofware metadata from as close to the
source as possible. This can be contrasted with other metadata efforts, also within CLARIN, to establish a more
centralised metadata store\footnote{Such as CLARIN's CMDI Component Registry;
\url{https://www.clarin.eu/content/component-metadata}} with more specialised and
research-oriented metadata. There is merit to both approaches as they serve distinct functions. But in order to best
serve everybody's different metadata needs, these bottom-up and top-down processes need to come to some kind of
synthesis, e.g. by having component registries do automatic harvesting of CodeMeta (or other) metadata, similar to what LaMachine
does.
